Ideal model selection is impossible

### 2014-10-21 Facticity/Facticity/paper.tex
allemaal stuff joh

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}{Definition}

@@

@@

\begin{lemma}
  Facticity is bounded.
\end{lemma}
\begin{proof}
Let $\bar\imath p$ be any two-part representation for the data $x$, i.e. $\phi_i(p)=x$. Then construct an alternative two-part representation $\bar vi^* p$, where $i^*$ is the shortest $\psi_u$-program for $i$ such that $\phi_v(i^* p)=\phi_{\psi_u(i^*)}(p) = \phi_i(p)=x$. We compare the lengths of these two representations. Note that $|i^*|=K^\psi(i)$. Therefore,
\[
|\bar\imath p|-|\bar v i^* p| = |\bar\imath|-|\bar v| - |i^*| = |\bar\imath|-|\bar v|-K(i).
\]
Since $|\bar\imath|$ is a computable codelength function, there must be an $i_0$ such that $|\bar\imath|-K(i)>|v|$ for all $i>i_0$ (see~\cite[Theorem~2.3.1]{li1993introduction}). From this $i$ onwards, the second representation (using $v$) will have a shorter code length, so any two-part representation $\bar ip$ cannot achieve the minimum in the definition of the facticity if $i>i_0$. Consequently, the facticity is bounded by $F^\phi(x)<|\overline{\imath_0}|$ for all $x$.
\end{proof}

>>>>

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{definition}{Definition}

@@

\newcommand{\argmin}{\mathop{\arg\min}}

@@

\[
b(i_0)=\min_{i\ge i_0}(L(i)-K(i))
\]

\begin{lemma}
  If $\min\{i\ge i_0:|\bar\imath|-K(i)\}$ is an unbounded function of $i_0$, then facticity is bounded.
\end{lemma}
\begin{proof}
Let $\bar\imath p$ be any two-part representation for the data $x$, i.e. $\phi_i(p)=x$. Then construct an alternative two-part representation $\bar vi^* p$, where $i^*$ is the shortest $\psi_u$-program for $i$ such that $\phi_v(i^* p)=\phi_{\psi_u(i^*)}(p) = \phi_i(p)=x$. We compare the lengths of these two representations. Note that $|i^*|=K^\psi(i)$. Therefore,
\[
|\bar\imath p|-|\bar v i^* p| = |\bar\imath|-|\bar v| - |i^*| = |\bar\imath|-|\bar v|-K(i).
\]
By assumption there must be an $i_0$ such that the above expression is positive for all $i>i_0$. From this $i$ onwards, the second representation (using $v$) will have a shorter code length, so $\bar\imath p$ cannot achieve the minimum in the definition of the facticity. Consequently, the facticity is bounded by $F^\phi(x)<|\overline{\imath_0}|$ for all $x$. 
\end{proof}
Note that the condition for this lemma applies to all known partial recursive prefix functions:

\begin{conjecture}
For any partial recursive prefix code length function $L$, the function
\[
\min\{L(i)-K(i):i\ge i_0\}
\]
is unbounded in $i_0$.
\end{conjecture}

\begin{definition}[Model complexity]
Given a model class $\M$, the model complexity of a data sequence $x$ is given by
\[G(x)=K(f^*)\qquad\tn{where}\qquad (f^*,p^*)=\argmin_{f,p:f\in\M, f(p)=x}(K(f)+|p|),
\]
where ties are broken by picking a model $f^*$ of minimal complexity.
\end{definition}

\begin{theorem}
There are enumerations $\phi_1,\phi_2,\ldots$ and $\psi_1,\psi_2,\ldots$ of the partial recursive (prefix) functions such that for every constant $c$ there is a binary sequence $x$ with
\[
|G^\phi(x)-G^\psi(x)|>c
\]
\end{theorem}

\begin{lemma}
There is an enumeration $\phi_1,\phi_2,\ldots$ of the partial recursive (prefix) functions, and a constant $c$ such that for all $n$ we have $G^\phi(x)<c$ for more than half of all strings $x$ of length $n$.
\end{lemma}

\begin{lemma}
There is an enumeration $\phi_1,\phi_2,\ldots$ of the partial recursive (prefix) functions such that for every constant $c$ there is an $n_0$ such that for all $n>n_0$ we have $G^\phi(x)>c$ for more than half of all strings $x$ of length $n$.
\end{lemma}

\begin{theorem}
There are enumerations $\phi_1,\phi_2,\ldots$ and $\psi_1,\psi_2,\ldots$ of the partial recursive (prefix) functions such that for every constant $c$ there is a binary sequence $x$ with
\[
|G^\phi(x)-G^\psi(x)|>c
\]
\end{theorem}
